\LARGE \bf How Children Perceive and Interact with a Robot that Behaves Unexpectedly -- The Domino Experiment

### 2014-09-29 paper.tex
Full proof-read. Still 10p. :-(

Presented is a study on the impact of unexpected robot behaviours on the
perception of a robot by children and their subsequent engagement in a playful interaction.
We propose an original analysis methodology which blends behavioural cues and
reported phenomenological perceptions into a compound \emph{index}.

@@

Different possibilities to foster engagement (both short- and long-term
engagement) in HRI have been explored, in particular with social robots. A lot
of research has moved toward creating sophisticated emotional models which cause
complex robot behaviour. \cite{leite_long-term_2013} studied the long-term
engagement of children with a chess playing robot that adapted its behaviour to
the children and showed empathy toward them. She found that empathetic robots
are more likely to engage users in the long-term and they proposed several
guidelines for designing such artificial companions. Some other work
\cite{bickmore_maintaining_2010,short_no_2010} has shown that there can be much
simpler ways to enhance engagement.

@@

Similarly, more looking at short-term engagement, \cite{short_no_2010} found
that a simple manipulation of the robot's behaviour can lead to greater
engagement. The authors let participants play several rounds of the
rock-paper-scissors game with the robot (the playfulness of the scenario seems
important). When the robot was cheating from time to time, participants tended
to ascribe intention to the robot what in turn led to greater engagement in how
they were interacting with the robot. The mechanism behind the cheating action
that the robot showed, was that participants did not expect this behaviour from
the robot. The authors suggest that \textit{``any deviation from expected
    operation is sufficient to create a greater degree of engagement in the
interaction.''} Moreover, \cite[p.~225]{short_no_2010} proposed that
\textit{``many interactions can be improved by the introduction of such simple
    behaviours, and that this should be exploited by designers in HRI. Bringing
    human and robot together to perform a simple, repetitive, familiar task and
then having the robot behave unexpectedly can increase engagement and mental
state attribution without complex behavioural or mechanical additions.''}.
In their model of the dynamics of anthropomorphism, Fink et
al.~\cite{fink2014dynamics} also suggest that \emph{disruptive behaviours}
may lead to increased anthropomorphic projections and possibly engagement.

@@

We present in this article a study that explores possibilities of sustaining
children's engagement with the \emph{Ranger}~\cite{mondada2014ranger} robot by
manipulating the robot's behaviour in such way that it appears
\textit{unexpected} to the children. We examine how different variations of
robot behaviour impact children's interaction with Ranger and their perception of
it (\eg in terms of attributing intention and cognitive abilities to the robot).

@@

This study also aims at applying to an actual study and assessing the usefulness
of the \emph{anthropomorphic index} proposed by Fink et
al.~\cite{fink2014dynamics} as a compound (behavioural and phenomenological)
metric for the attribution of human-like features to robot. We are especially
interested in investigating how this indicator predicts and/or reflects actual
\emph{engagement} into the interaction.

\subsection{Research Questions and Hypotheses}

In the Domino Study, we analyze child-robot interaction with a robot that shows
unexpected behaviour. In a playful scenario which was set up in a laboratory
environment, 26 children aged 4-5 years were assembling a domino game together.
Each group consisted of two children and the \emph{Ranger} robot, which was used
to transport domino tiles between the two children.

@@

    \item The robot gets \textbf{lost}: When called by the child to come over,
        the robot goes wrong, without any observable reason, and remains at the
        wrong location. We expect this to be perceived as a mechanical
        malfunction (a bug or system error which causes the robot to not work
        correctly), and hypothesize decreased attributions of human-likeness to
        the robot.

@@

    \item The robot makes a \textbf{mistake}: When called by the child to come
        over, the robot goes wrong but recognizes its mistake and repairs. We
        expect this to be perceived (explicitly) as \textit{``to err is
        human''}, and (implicitly) as the robot being endowed with a certain
        level of introspective capabilities (it was able to recognize its own
        error). In this condition, we assume increased attributions of
        human-likeness to the robot.\footnote{The attribution of human-likeness
            to a robot is hereafter labeled as \textit{anthropomorphic
        projections}.}

@@

Our research questions deal with both children's observable behaviour and their
perception of the robot. We would also like to explore the relation between
these two aspects, related to the attribution of human-like characteristics to
the robot. The motivation behind this is to find out whether
\textit{anthropomorphism} can not only be measured as a specific type of
perception but also as an observable behaviour in the interaction itself. Thus,
we would like to bridge things together and, relying on the concept of an
\textbf{anthropomorphism index}, build a
compound metric that considers both children's perception and interaction
aspects (qualitatively). Previous work suggests that a social relation to a
robot (we view anthropomorphism as a specific type of social relation) reflects
an increased engagement and can be effective in sustaining interaction.
Consequently, we formulate a third hypotesis:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scenario and Research Methodology}

\subsection{The Domino scenario}

@@

The interaction scenario consisted in two children who play domino together,
with the help of the robotic box Ranger. The scenario setup is displayed in
Figure~\ref{fig:domino-setup}. The challenge is that the tiles of the domino are
distributed in the room, hidden behind three beanbags, and while one child --
\textit{the searcher} -- searches for the right tile, the other child --
\textit{the receiver} -- is asked to stay in a play tent. There is a ``river''
(play carpet) in between the two children that we told them they cannot cross
and therefore they need the robot to transport the domino tiles between them.

The Ranger~\cite{mondada2014ranger} is a wheeled box (27 x 37 x 37 cm) with
partial wooden surface. Being remote controlled, it can move around a flat
surface, move its eyes and eyebrows, display colors (LEDs) and light patterns,
and play sounds through Bluetooth speakers.  The robot was controlled by a human
wizard, who was in the same room (see Figure~\ref{fig:domino-setup}). This was
to ensure that the wizard could see children's interaction and hear their
commands to the robot without any delay.  Only one group asked at the beginning
of the experiment if the wizard was the one actually controlling the robot. This
did not seem to subsequently influence their behaviour.

We used a self-made domino consisting of 10 wooden tiles (10~x~20~x~1.5~cm) with
pictures of cartoon farm animals: cow, sheep, hen, donkey, duck, pig, rabbit.
The pictures were taken from a commercially available domino game adapted to the
age of the children (produced by Djeco, advised age 3+).

To start the game (divided in several \emph{runs} that correspond each to the
delivery and assembling of one domino tile), there is already one domino tile in
front of the tent, where the receiver child stays and assembles the domino
chain. The receiver child asks the searcher child for a specific tile, \eg a
tile with a donkey, the searcher searches for the respective tile, sits down on
the next beanbag, and asks the robot to come over.  The Ranger robot is first
located next to the tent, then, when called by the searcher it starts moving,
crosses the river carpet, and comes over to the searcher on the beanbag. The
searcher child puts the domino tile into the robotic box, and the robot then
goes back to the receiver child in the tent. Then, the receiver takes out the
domino tile from the robot, and puts the two tiles together. The first
\emph{run} is over, and a new \emph{run} starts, when the receiver asks the
searcher for the next domino tile.

@@

Special attention has been paid to the distribution of misbehaving runs during
the study. The five first correct runs aims at setting the children expectation
regarding a consistent robot behaviour. When the robot then exhibits an
unexpected behaviour, children are likely to be positively surprised. A similar
effect has been observed in a study with a robot that was cheating from time to
time~\cite{short_no_2010}.  Besides, to prevent the robot misbehaviours to be
immediately interpreted as failures, we introduce these misbehaviour neither at
the very beginning nor at very end of the interaction
(\cite{desai_effects_2012,desai_impact_2013} show that early and late robot
failures negatively impact trust). \cite{short_no_2010} also adopted a similar
pattern when introducing unexpected robot behaviour. In their study with a robot
that cheated, from the 20 rounds of the rock-paper-scissors game, the robot
cheated three times in the middle: on the 4th, 8th and 15th round. The authors
adjusted the interaction length so that the spacing and timing of the cheating
was preserved.

@@

The \emph{correct} robot behaviour (baseline behaviour) consists in the
following states:

    \item \emph{Domino tile put in robot:} Ranger makes a ``rewarding'' sound
        and shows a green light pattern,

    \item \emph{Domino tile removed from robot:} Ranger makes an ``emptying''
        sound and shows a green light pattern,

    \item \emph{Robot is called by one of the children:} When one of the
    children called the robot saying something like \textit{``Robot come
    here!''}, the robot starts moving toward the child (Ranger does not react to
    any other verbal commands).

    \item \emph{Robot reaches one of the children:} When in front of one of the
    children who should either put or remove a domino tile, Ranger stops and
    shows a pulsating yellow light pattern. If no reaction, Ranger also makes a
    wiggle-like move.

During the misbehaving runs, the robot's behaviour is manipulated in three
possible ways (Figure~\ref{fig:domino-misbehaviour}):

@@

We have obtained and analysed two types of data. The \textbf{children's
behaviours toward the robot} (\ie the child-robot interactions) have been
captured in the video recordings by annotating a set of actions, described
below. The children's \textbf{phenomenological perception of the robot} has been
captured through audio-recorded semi-structured interviews which took place
between run \emph{1.5} and \emph{2.1} and at the end of the 14 runs.

@@

    \item \textbf{Explore (ex)}: when children actively try to find out what the
        robot is doing (\eg by looking under the box); attentively watch or
        observe the robot (\eg attentively waiting for the box to show a
        reaction); experiment with the robot to figure out how it works (\eg put
        hands in front or inside of the box to see what happens);

    \item \textbf{Misuse (mis)}: when children kick the robot, poke it in its
        ``eye'', try to climb on or inside the box, drive / push the robot
        around, stop the robot's wheels with a foot;

    \item \textbf{Put domino (put)}: when a domino tile is put inside the box;

@@

    \item \textbf{Gesture (ges)}: when gestures are used to communicate /
        interact with the robot (\eg pointing gestures, waving at the robot);

    \item \textbf{Touch (touch)}: when the box is touched (\eg petted or
        caressed);

    \item \textbf{Show (show)}: when a child shows something to the robot (\eg
        by holding a domino tile in front of its eyes);

    \item \textbf{Call (call)}: when a child calls the robot to come over;

    \item \textbf{Talk (talk)}: when a child directly talks to the robot (using
        direct speech) besides calling the robot;

    \item \textbf{Look (look)}:	when a child looks at the experimenter due to
        confusion caused by the robot; look is not coded when the experimenter
        asks a question to the child;

@@

The actions \textit{put, remove} and \textit{call} were ``requested'' actions
because the scenario required them and children were asked to carry them out.
Hence, these actions are not relevant to be analyzed quantitatively. The other
actions \textit{explore, misuse, gesture, touch, show, talk} and \textit{look}
are spontaneous actions that were not requested but arose in the interaction.
These actions reflect engagement, and are the one we considered during the
analysis. Figure~\ref{fig:domino-time-active} shows the distribution of these
actions over the different runs, in any condition.

@@

\begin{table}[h!t]
\centering
\footnotesize
\begin{tabular}{p{0.9\linewidth}}
    \toprule
    \textbf{Construct \emph{expectations}} \\
    \midrule

@@

    %%
    \toprule
    \textbf{Construct \emph{impression}} \\
    \midrule

@@

In designing our interview script and selecting relevant questions, we took
inspiration from previous work on child-robot interaction and children's
perception of robots
\cite{kahn_jr._robotic_2006,weiss_i_2009,leite_influence_2013}. For instance, we
applied and adapted some of the \emph{constructs} and example questions from the
questionnaires used in \cite{kahn_jr._robotic_2006} and \cite{weiss_i_2009}. A
\emph{construct} addresses a specific factor (topic) that can be measured by
several questions. For instance, the construct ``cognitive connections'' (using
Flavell's terminology~\cite{flavell1988development}) considers the robot's
ability to hear and to see (perceptual skills), as attributed by the children.
The construct ``moral standing'' and the respective question was taken from
\cite{kahn_jr._robotic_2006}.\footnote{According to
\cite{kahn_jr._robotic_2006}, \textit{moral} refers to considerations based
on an artifact's physical or psychological welfare, and virtue (whether the
artifact deserves care). An attribution of moral standing reflects, for
instance, that the robot engenders moral regard, is morally responsible,
blameworthy, has rights or deserves respect.}			Similarly, we grouped
questions according to specific constructs that they evaluate (see
Table~\ref{tab:domino-questions}).

With several recurring questions in the first and second interview, we wanted to
see the differences in children's perception of the correctly behaving and
unexpectedly behaving robot. We planned to use these two interviews as a
within-subject measurement, however, this did not work out very well because
children's responses were not always accurate, not comparable one by one, and
children did not always give an answer.

@@

\subsubsection{Engagement and anthropomorphism index}

There are different possibilities to measure engagement in HRI, depending on the
specific research question and context. Metrics to measure behavioural engagement
in the interaction include, for instance, conversation analysis (\eg used in
\cite{short_no_2010}) or general attention analysis. These can be studied by
analyzing interaction videos in terms of head movement, eye tracking and gesture
/ body movement analysis (\eg in \cite{sidner_explorations_2005}).
Post-measurements to measure emotional engagement can be questionnaires that try
to assess constructs like the perceived presence and involvement in the
interaction. An example is the \textit{Interactive Experience Questionnaire}
(originally developed by \cite{lombard_measuring_2000}) of which adapted
versions were used in
\cite{kidd_effect_2004,bainbridge_effect_2008,short_no_2010}.

A mix of several methods has been used in a long-term interaction study with 8-9
year old children, \cite{leite_long-term_2013}. The author measured engagement
through video observations (by analyzing the amount of time that children spent
looking at the robot), interviews, and questionnaires. The interviews were
semi-structured, containing initial yes-or-no questions followed by open-ended
questions that allowed children to justify and elaborate their answers. We have
chosen a similar approach that considers both children's behavioural and
emotional engagement.

With 4-5 year old children, however, we cannot use rating scales to ask them
about constructs as abstract as social presence or how much they felt involved
in the interaction. Therefore we have attempted to combine \emph{interaction}
data with the results of the semi-structured interviews.  Several of the
aforementioned coded actions can reflect engagement: \textit{explore, misuse,
gesture, touch, show,} and \textit{talk}. Also, \textit{look} at the
experimenter can be considered as engagement in the interaction: We assume that
children look at the experimenter when they are surprised and seek for help.
This behaviour reflects that they want to make sense of the robot's behaviour and
that they notice it as ``strange''. Further, we can take into account how they
refer to Ranger and describe the robot (as well as their experience) in the
interviews.  Specifically, we used the constructs \textit{ascribe intention,
    ascribe cognitive abilities, ascribe emotional state}, and \textit{ascribe
    moral standing} (Table~\ref{tab:domino-questions}) to assess how far they
    anthropomorphize the robot.

To account for the fact that anthropomorphism arises in an interaction, we try
to bring both children's perception of the robot (post-measurement) and their
behaviour toward it (in-the-moment measurement) together. By doing so, we would
like to obtain a qualitative \textbf{anthropomorphism
index}~\cite{fink2014dynamics}. We quantify the index by attributing points for each
anthropomorphic \emph{perception} of the robot (max.  13 points) and for specific kinds
of human-like \emph{behaviour} toward the robot (max.  3 points). The index included the
following aspects:\\

@@

There are certainly limitations to this scoring scheme. For instance, the
balance between perception and behaviour aspects is questionable (13 to 3
points). Also, we did not consistently assign 1 point to each item, but assigned
points between 0.5 and 2 points. We did this because we found that different
items reflect a higher level of anthropomorphic perception of the robot than
others (for instance, ascribing the ability to see and hear was suggested by our
study setup, and we cannot be sure that it really reflects anthropomorphism).
There are possibilities of refinement of this scoring scheme. For instance, a
more systematic weighting procedure and the identification of all relevant
factors would be an interesting future achievement. We could introduce a formula
to calculate the anthropomorphism index \textit{$\alpha$} by taking the sum of
the various items ($x_1, x_2, x_3...$) and different weights ($a, b, c ...$) to
balance them: $\alpha$ \textit{=} $a x_1$ \textit{+} $b x_2$ \textit{+} $c x_3
...$~. Machine learning techniques could be used to optimize such a calculation
of the anthropomorphism index.

@@

In terms of \textbf{engagement}, the novelty of the robot certainly plays a role
here. As mentioned before, the findings here do not directly address the issues
of long-term usage but concern short-term \textit{engagement}, which is a
pre-requisite for long-term usage.

The huge proportion of \textit{explore} actions (52~\% of all coded actions)
already suggests that children were generally engaged in the interaction, and
most of the time carefully observed what the robot did. When considering the
actions \textit{explore, gesture, look, misuse, show, talk}, and \textit{touch}
as \textbf{engagement actions}, 1699 of the 2354 actions reflected engagement
(72~\%). This indicates that overall children were very engaged in the
interaction with the robot. Furthermore, data suggests that the robot in the
\textit{mistake} and \textit{lost} condition were more engaging for children
(each 75~\% of the actions reflected engagement) than the robot in the
\textit{disobey} condition (66~\% reflected engagement). The statistical
analysis supports this: An ANOVA indicates that there was a significant effect
of robot behaviour on the actions \textit{explore} (F(2,23)=11.31, p<.001) and
\textit{look} (F(2,23)=4.6, p=.021). Post-hoc comparisons using the Tukey's test
indicate that the mean score for \textit{explore} in the \textit{disobey}
condition is significantly different from the score in the \textit{mistake} and
\textit{lost} condition. This suggests that children explored the disobeying
robot less\footnote{It is surprising that the disobeying robot was explored
less. One may expect that the disobeying robot might better attract
children's attention because it faces the searcher child and uses fairly
strong audio and light cues as compared to the other behaviours.}. For
\textit{look}, the only significant difference is found between the mean
scores of \textit{lost} and \textit{disobey}. Data suggests that when the
robot is lost, children look more often at the experimenter than when the
robot disobeys. The other actions were not found to differ significantly
between the robot behaviour conditions.

Overall, we found a significant difference between the average of engagement
actions carried out during the first 7 runs (correct robot behaviour) and during
the second 7 runs, when the robot behaved unexpectedly (F(1,36)=5.1, p=.03). In
all three conditions, children carried out more engagement actions with the
unexpectedly behaving robot. No interaction effect was found between the two
phases of interaction (expected / unexpected) and condition (F(2,36)=1.2,
p=.31). In general, this finding supports our first hypothesis: children show
more engagement toward a robot that behaves unexpectedly from time to time
compared to a robot that always behaves correctly.

@@

But to what extent is the human-like interaction related to perceiving the robot
as human-like? Do children who interact a lot and are probably more engaged with
the robot also perceive the robot as more human-like? This was what we
hypothesized. Data suggests the opposite, however. As shown in
Figure~\ref{fig:domino-anthropo-interaction}, we found a significant
\textit{negative} correlation between the count of engagement actions (per
group) and the qualitative anthropomorphism index (r(11)=-0.56, p=.05).

>>>>

Presented is a study on the impact of unexpected robot behaviours on the
perception of a robot by children and their subsequent engagement in a playful interaction.
We propose an original analysis methodology which blends behavioural cues and
reported phenomenological perceptions into a compound index.

@@

Different possibilities to foster engagement (both short- and long-term
engagement) in HRI have been explored, in particular with social robots. A lot
of research has moved toward creating sophisticated emotional models which cause
complex robot behaviour. \cite{leite_long-term_2013} studied the long-term
engagement of children with a chess playing robot that adapted its behaviour to
the children and showed empathy toward them. She found that empathetic robots
are more likely to engage users in the long-term and they proposed several
guidelines for designing such artificial companions. Other works
\cite{bickmore_maintaining_2010,short_no_2010} have shown that simpler ways to
enhance engagement may as well be effective.

@@

Similarly, looking at short-term engagement, \cite{short_no_2010} found that a
simple manipulation of the robot's behaviour can lead to greater engagement. The
authors let participants play several rounds of the rock-paper-scissors game
with the robot (the playfulness of the scenario seems important). When the robot
was cheating from time to time, participants tended to ascribe intention to the
robot what in turn led to greater engagement in how they were interacting with
the robot. The authors observe that \textit{``any deviation from expected
operation is sufficient to create a greater degree of engagement in the
interaction.''} Besides, they suggest that \textit{``many interactions can be
improved by the introduction of such simple behaviours, and that this should
be exploited by designers in HRI. Bringing human and robot together to
perform a simple, repetitive, familiar task and then having the robot behave
unexpectedly can increase engagement and mental state attribution without
complex behavioural or mechanical additions.''}.  Along those lines, Fink
\etal\cite{fink2014dynamics} also suggest in their model of the dynamics of
anthropomorphism, that \emph{disruptive behaviours} may lead to increased
anthropomorphic projections and possibly engagement.

@@

We present in this article a study that explores possibilities of sustaining
children's engagement with the \emph{Ranger} robot~\cite{mondada2014ranger} by
manipulating the robot's behaviour in such way that it appears
\textit{unexpected} to the children. We examine how different variations of
robot behaviour impact children's interaction with Ranger and their perception of
it (in terms of attributing intention and cognitive abilities to the robot).

@@

This study also aims at assessing the effectiveness of the \emph{anthropomorphic
index} proposed by Fink \etal\cite{fink2014dynamics} as a compound (behavioural
and phenomenological) metric for the attribution of human-like features to
robot. We are especially interested in investigating how this indicator predicts
and/or reflects actual \emph{engagement} into the interaction.

\subsection{Design and Hypotheses}

In the Domino Study, we analyze child-robot interaction with a robot that shows
unexpected behaviour. In a playful scenario which was set up in a laboratory
environment, 26 children aged 4-5 years (MD=4.46) were assembling a domino game together.
Each group consisted of two children and the \emph{Ranger} robot, which was used
to transport domino tiles between the two children.

@@

    \item The robot gets \textbf{lost}: When called by the child to come over,
        the robot goes wrong, without any observable reason, and remains at the
        wrong location. We expect this to be perceived as a mechanical
        malfunction (a bug or system error which causes the robot to not work
        correctly), and hypothesize decreased attributions of human-likeness to
        the robot.\footnote{We use interchangeably \textit{attribution of
            human-likeness} and \textit{anthropomorphic projections} in the
        remaining of the article.}

@@

    \item The robot makes a \textbf{mistake}: When called by the child to come
        over, the robot goes wrong but recognizes its mistake and repairs. We
        expect this to be perceived (explicitly) as \textit{``to err is
        human''}, and (implicitly) as the robot being endowed with a certain
        level of introspective capabilities (it was able to recognize its own
        error). In this condition, we assume increased attributions of
        human-likeness to the robot.

@@

Our research questions deal with both children's observable behaviour and their
perception of the robot. We would also like to explore the relation between
these two aspects, related to the attribution of human-like characteristics to
the robot. The motivation behind this is to find out whether
\textit{anthropomorphism} can not only be measured as a specific type of
perception but also as an observable behaviour in the interaction itself. Thus,
we would like to bridge things together and, relying on the concept of
anthropomorphism index~\cite{fink2014dynamics}, build a compound metric that
considers both children's perception and interaction aspects (qualitatively).
Previous work suggests that a social relation to a robot (we view
anthropomorphism as a specific type of social relation) reflects an increased
engagement and can be effective in sustaining interaction.  Consequently, we
formulate a third hypothesis:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Domino study}

\subsection{Scenario and course of the study}

@@

The interaction scenario consisted in two children who play domino together,
with the help of the robotic box Ranger. The scenario setup is displayed in
Figure~\ref{fig:domino-setup}. The challenge is that the tiles of the domino are
distributed in the room, hidden behind three beanbags, and while one child --
\textit{the searcher} -- searches for the right tile, the other child --
\textit{the receiver} -- is asked to stay in a play tent. There is a ``river''
(playground) in between the two children that we told them they cannot cross
and therefore they need the robot to transport the domino tiles between them.

The Ranger~\cite{mondada2014ranger} is a wheeled box (27 x 37 x 37 cm) with
partial wooden surface. It is remotely controlled, it can move around a flat
surface, move its eyes and eyebrows, display colors (LEDs) and light patterns,
and play sounds through Bluetooth speakers.  The robot was controlled by a human
wizard, who was in the same room (see Figure~\ref{fig:domino-setup} -- only one
group asked at the beginning of the experiment if the wizard was the one
actually controlling the robot. This did not seem to subsequently influence
their behaviour).

We used a self-made domino consisting of 10 wooden tiles (10~x~20~x~1.5~cm) with
pictures of cartoon farm animals (taken from a commercially available domino
game manufactured by Djeco and adapted to the age of the children).

The game (divided in several \emph{runs} that correspond each to the delivery
and assembling of one domino tile) starts with one domino tile in front of the
tent, where the receiver child stays and assembles the domino chain. The
receiver child asks the searcher child for a specific tile, \eg a tile with a
donkey, the searcher searches for the respective tile, sits down on the next
beanbag, and asks the robot to come over.  The Ranger robot is first located
next to the tent, then, when called by the searcher (\textit{``Robot, come
here!''}), it starts moving, crosses the river carpet, and comes over to the
searcher on the beanbag. The searcher child puts the domino tile into the
robotic box, and the robot then goes back to the receiver child in the tent.
Then, the receiver takes out the domino tile from the robot, and puts the two
tiles together. The first \emph{run} is over, and a new \emph{run} starts, when
the receiver asks the searcher for the next domino tile.

@@

Special attention has been paid to the distribution of misbehaving runs during
the study. The five first correct runs aims at setting the children expectation
regarding a consistent robot behaviour. When the robot then exhibits an
unexpected behaviour, children are likely to be positively surprised (a similar
effect has been observed in a study with a robot that was cheating from time to
time~\cite{short_no_2010}).  Besides, to prevent the robot misbehaviours to be
immediately interpreted as failures, we introduce these misbehaviour neither at
the very beginning nor at very end of the interaction
(\cite{desai_effects_2012,desai_impact_2013} show that early and late robot
failures negatively impact trust. \cite{short_no_2010} also adopted a similar
pattern when introducing unexpected robot behaviour).

@@

%The \emph{correct} robot behaviour (baseline behaviour) consists in the
%following states:
%
%\begin{itemize}
%
%    \item \emph{Domino tile put in robot:} Ranger makes a ``rewarding'' sound
%        and shows a green light pattern,
%
%    \item \emph{Domino tile removed from robot:} Ranger makes an ``emptying''
%        sound and shows a green light pattern,
%
%    \item \emph{Robot is called by one of the children:} When one of the
%    children called the robot saying something like \textit{``Robot come
%    here!''}, the robot starts moving toward the child (Ranger does not react to
%    any other verbal commands).
%
%    \item \emph{Robot reaches one of the children:} When in front of one of the
%    children who should either put or remove a domino tile, Ranger stops and
%    shows a pulsating yellow light pattern. If no reaction, Ranger also makes a
%    wiggle-like move.
%
%\end{itemize}
%
During the misbehaving runs, the robot's behaviour is manipulated in three
possible ways (Figure~\ref{fig:domino-misbehaviour}):

@@

We have obtained and analysed two types of data. The \textbf{children's
behaviours toward the robot} (\ie the child-robot interactions) have been
captured in the video recordings by annotating a set of actions, described
below. The children's \textbf{phenomenological perception of the robot} has been
captured through two audio-recorded semi-structured interviews which took place
between run \emph{1.5} and \emph{2.1} and at the end of the 14 runs.

@@

    \item \textbf{Explore}: when children actively try to find out what the
        robot is doing (\eg by looking under the box); attentively watch or
        observe the robot (\eg attentively waiting for the box to show a
        reaction); experiment with the robot to figure out how it works (\eg put
        hands in front or inside of the box to see what happens);

    \item \textbf{Misuse}: when children kick the robot, poke it in its
        ``eye'', try to climb on or inside the box, drive / push the robot
        around, stop the robot's wheels with a foot;

    \item \textbf{Put domino}: when a domino tile is put inside the box;

@@

    \item \textbf{Gesture}: when gestures are used to communicate /
        interact with the robot (\eg pointing gestures, waving at the robot);

    \item \textbf{Touch}: when the box is touched (\eg petted or
        caressed);

    \item \textbf{Show}: when a child shows something to the robot (\eg
        by holding a domino tile in front of its eyes);

    \item \textbf{Call}: when a child calls the robot to come over;

    \item \textbf{Talk}: when a child directly talks to the robot (using
        direct speech) besides calling the robot;

    \item \textbf{Look}: when a child looks at the experimenter due to
        confusion caused by the robot; look is not coded when the experimenter
        asks a question to the child;

@@

The actions \textit{put, remove} and \textit{call} were ``requested'' actions
that the children had to perform in the scenario. On the contrary,
\textit{explore, misuse, gesture, touch, show, talk} and \textit{look} are
spontaneous actions that were not requested but arose in the interaction.  These
actions reflect engagement, and are the one we considered during the analysis.
Figure~\ref{fig:domino-time-active} shows the distribution of these actions over
the different runs, in summed over the three condition.

@@

\begin{table}[h!t]
\centering
\footnotesize
\begin{tabular}{p{0.9\linewidth}}
    \toprule
    \textbf{Expectations} \\
    \midrule

@@

    %%
    \toprule
    \textbf{Impression} \\
    \midrule

@@

In designing our interview script and selecting relevant questions, we took
inspiration from previous work on child-robot interaction and children's
perception of robots
\cite{kahn_jr._robotic_2006,weiss_i_2009,leite_influence_2013}. For instance, we
applied and adapted some of the \emph{constructs} and example questions from the
questionnaires used in \cite{kahn_jr._robotic_2006} and \cite{weiss_i_2009}. A
\emph{construct} addresses a specific factor (topic) that can be measured by
several questions. For instance, the construct ``cognitive connections'' (using
Flavell's terminology~\cite{flavell1988development}) considers the robot's
ability to hear and to see (perceptual skills), as attributed by the children.
The construct ``moral standing'' and the related question was taken from
\cite{kahn_jr._robotic_2006}.\footnote{According to
\cite{kahn_jr._robotic_2006}, \textit{moral} refers to considerations based
on an artifact's physical or psychological welfare, and virtue (whether the
artifact deserves care). An attribution of moral standing reflects, for
instance, that the robot engenders moral regard, is morally responsible,
blameworthy, has rights or deserves respect.} Similarly, we grouped
questions according to specific constructs that they evaluate (see
Table~\ref{tab:domino-questions}).

With several recurring questions in the first and second interview, we wanted to
see the differences in children's perception of the correctly behaving and
unexpectedly behaving robot. We planned to use these two interviews as a
within-subject measurement, however, this did not fully work out because
children's responses were not always accurate, not comparable one by one, and
children did not always give an answer.

@@

\subsubsection{Engagement}

%There are different possibilities to measure engagement in HRI, depending on the
%specific research question and context. Metrics to measure behavioural engagement
%in the interaction include, for instance, conversation analysis (\eg used in
%\cite{short_no_2010}) or general attention analysis. These can be studied by
%analyzing interaction videos in terms of head movement, eye tracking and gesture
%/ body movement analysis (\eg in \cite{sidner_explorations_2005}).
%Post-measurements to measure emotional engagement can be questionnaires that try
%to assess constructs like the perceived presence and involvement in the
%interaction. An example is the \textit{Interactive Experience Questionnaire}
%(originally developed by \cite{lombard_measuring_2000}) of which adapted
%versions were used in
%\cite{kidd_effect_2004,bainbridge_effect_2008,short_no_2010}.
%
%A mix of several methods has been used in a long-term interaction study with 8-9
%year old children, \cite{leite_long-term_2013}. The author measured engagement
%through video observations (by analyzing the amount of time that children spent
%looking at the robot), interviews, and questionnaires. The interviews were
%semi-structured, containing initial yes-or-no questions followed by open-ended
%questions that allowed children to justify and elaborate their answers. We have
%chosen a similar approach that considers both children's behavioural and
%emotional engagement.
%
%With 4-5 year old children, however, we cannot use rating scales to ask them
%about constructs as abstract as social presence or how much they felt involved
%in the interaction.

We have measured engagement by combining \emph{interaction}
data with the results of the semi-structured interviews.  Several of the
aforementioned coded actions can reflect engagement: \textit{explore, misuse,
gesture, touch, show,} and \textit{talk}. Also, \textit{look} at the
experimenter can be considered as engagement in the interaction: We assume that
children look at the experimenter when they are surprised and seek for help.
This behaviour reflects that they want to make sense of the robot's behaviour and
that they notice it as ``strange''. Further, we can take into account how they
refer to Ranger and describe the robot (as well as their experience) in the
interviews.  Specifically, we used the constructs \textit{ascribe intention,
ascribe cognitive abilities, ascribe emotional state}, and \textit{ascribe
moral standing} (Table~\ref{tab:domino-questions}) to assess how far they
engaged with the robot.

To account for the fact that anthropomorphism arises in an interaction, we tried
to bring both children's perception of the robot (post-measurement) and their
behaviour toward it (in-the-moment measurement) together. We did so by building
a qualitative anthropomorphism index~\cite{fink2014dynamics}. We quantify the
index by attributing points for each anthropomorphic \emph{perception} of the
robot (max.  13 points) and for specific kinds of human-like \emph{behaviour}
toward the robot (max.  3 points). The index included the following aspects:\\

@@

There are certainly limitations to this scoring scheme. For instance, the
balance between perception and behaviour aspects is questionable (13 to 3
points). Also, we did not consistently assign 1 point to each item, but assigned
points between 0.5 and 2 points. We did this because we found that different
items reflect a higher level of anthropomorphic perception of the robot than
others (for instance, ascribing the ability to see and hear was suggested by our
study setup, and we cannot be sure that it really reflects anthropomorphism).

@@

In terms of \textbf{engagement}, the novelty of the robot certainly plays a role
here. The findings here do not directly address the issues
of long-term usage but concern short-term \textit{engagement}, which is a
pre-requisite for long-term usage.

The huge proportion of \textit{explore} actions (52~\% of all coded actions)
already suggests that children were generally engaged in the interaction, and
most of the time carefully observed what the robot did. When considering the
actions \textit{explore, gesture, look, misuse, show, talk}, and \textit{touch}
as \textbf{engagement actions}, 1699 of the 2354 actions reflected engagement
(72~\%). This indicates that overall children were very engaged in the
interaction with the robot. Furthermore, data suggests that the robot in the
\textit{mistake} and \textit{lost} condition were more engaging for children
(each 75~\% of the actions reflected engagement) than the robot in the
\textit{disobey} condition (66~\% reflected engagement). The statistical
analysis supports this: An ANOVA indicates that there was a significant effect
of robot behaviour on the actions \textit{explore} (F(2,23)=11.31, p<.001) and
\textit{look} (F(2,23)=4.6, p=.021). Post-hoc comparisons using the Tukey's test
indicate that the mean score for \textit{explore} in the \textit{disobey}
condition is significantly different from the score in the \textit{mistake} and
\textit{lost} condition. This suggests that children explored the disobeying
robot less\footnote{It is surprising that the disobeying robot was explored
less. One may expect that the disobeying robot might better attract
children's attention because it faces the searcher child and uses fairly
strong audio and light cues as compared to the other behaviours.}.

Overall, we found a significant difference between the average of engagement
actions carried out during the first 7 runs (correct robot behaviour) and during
the second 7 runs, when the robot behaved unexpectedly (F(1,36)=5.1, p=.03). In
all three conditions, children carried out more engagement actions with the
unexpectedly behaving robot (Figure~\ref{fig:domino-time-active}). No
interaction effect was found between the two phases of interaction (expected /
unexpected) and condition (F(2,36)=1.2, p=.31). In general, this finding
supports our first hypothesis: children show more engagement toward a robot that
behaves unexpectedly from time to time compared to a robot that always behaves
correctly.

@@

We hypothesized that children who interact a lot and are probably more engaged
with the robot also perceive the robot as more human-like. Data suggests the
opposite, however. As shown in Figure~\ref{fig:domino-anthropo-interaction}, we
found a significant \textit{negative} correlation between the count of
engagement actions (per group) and the qualitative anthropomorphism index
(r(11)=-0.56, p=.05).
